D3064R0: How to Avoid OOTA Without Really Trying

### 2024-03-05 ootavacuous.tex
"Illustrative Litmus Tests": Reorder in decreasing importance
This commit pulls to the beginning the most important, surprising,
and illuminating sections of this appendix.

Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
"Illustrative Litmus Tests": Update lead-in paragraph
Document the rationale for the section ordering.

Signed-off-by: Paul E. McKenney <paulmck@kernel.org>
abstract: Rework sdep challenges
Suggested-by: Alan Stern <stern@rowland.harvard.edu>
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

\begin{abstract}
	The out-of-thin-air (OOTA) properties
	of \co{memory_order_relaxed} in the C++ memory model
	have caused considerable consternation over the years.
	Attempts to create memory models that rule out OOTA behaviors
	have been either non-executable, complex, or unloved by C++
	implementers
	(and this paper describes some of the challenges of coming to
	grips with the key notion of semantic dependency).
	But at the same time, we know of no instances of OOTA behavior
	in real C++ implementations.

	We focus on shared-memory programs and
	C++ implementations based on traditional compilers and
	computing hardware, including CPUs and GPGPUs.
	This context permits us to consider the detailed relations
	between source code and machine code required by the restricted
	nature of volatile atomic accesses.
	The OOTA problem is much more tractable at the hardware level
	than at the source level, thanks to existing formal hardware
	models, and
	we show that the constraints in these models prevent OOTA cycles
	involving only volatile atomics from occurring in
	undefined-behavior-free C++ programs (under these implementations).
	We also extend this work to nonvolatile atomics by defining
	``quasi-volatile'' behavior that we expect C++ implementations
	will adhere to if they perform single-thread analysis.
\end{abstract}

@@

These litmus tests helped illuminate important aspects of the problem
of defining sdep and identifying OOTA cycles.
This appendix evaluates them using the \co{herd7} tool and also using
manual analysis.

>>>>

\begin{abstract}
	The out-of-thin-air (OOTA) properties
	of \co{memory_order_relaxed} in the C++ memory model
	have caused considerable consternation over the years.
	Attempts to create memory models that rule out OOTA behaviors
	have been either non-executable, complex, or unloved by C++
	implementers.
	But at the same time, we know of no instances of OOTA behavior
	in real C++ implementations.

	We focus on shared-memory programs and
	C++ implementations based on traditional compilers and
	computing hardware, including CPUs and GPGPUs.
	This context permits us to consider the detailed relations
	between source code and machine code required by the restricted
	nature of volatile atomic accesses.
	The OOTA problem and the related challenge of coming to grips
	with semantic dependency are much more tractable at the hardware
	level than at the source level, thanks to existing formal
	hardware models.
	We show that these models' constraints prevent OOTA cycles
	that involve only volatile atomics from occurring in
	undefined-behavior-free C++ programs (under these implementations).
	We also extend this work to nonvolatile atomics by defining
	``quasi-volatile'' behavior that we expect C++ implementations
	will adhere to if they perform single-thread analysis.
\end{abstract}

@@

These litmus tests helped illuminate important aspects of the problem
of defining sdep and identifying OOTA cycles.
This appendix presents these tests in roughly decreasing order of
importance, surprise, and illumination.
It evaluates them using the \co{herd7} tool and also using manual
analysis.
